---
title: "Final Project"
author: "Angela Tan"
date: "2022-11-29"
output:
  pdf_document: default
  html_document: default
---

The study I am using: 

https://www.science.org/doi/10.1126/sciadv.aba9221 

```{r message=FALSE}
# Load required packages
library(ggplot2)
library(plyr)
library(dplyr)
#library(reshape)
library(igraph)
library(ggdag)
library(tidyverse)
library(igraph)
library(CINNA)

source(file="FormNets_OnlineMaterial.R")

```

```{r}
#provided data files
att <- read.csv("Attributes.csv")  # file of participant attributes
att$PPID <- as.character(att$PPID)
snat1 <- read.csv("SNA_T1.csv")      # time 1 network data
snat2 <- read.csv("SNA_T2.csv")      # time 1 network data

# final datasets after running FromNets_OnlineMaterial.R - quantified social network strengths 
# the 290 students who completed the intervention 
int290 <- read.csv("int290.csv")
int290$PPID <- as.character(int290$PPID)

# the 226 students who completed 
complet226 <- read.csv("complet226.csv") 
complet226$PPID <- as.character(complet226$PPID)
```

# Participant descriptives

Condition breakdown among all 290 students who completed the intervention:

```{r}
table(int290$Intervention)

ggplot(data = att, aes( x = factor(Intervention), fill = Intervention) ) +
  geom_bar() +
  geom_text(stat='count', aes(label=..count..), vjust=0)
```
Of the 328 students who consented to participate, 290 students (88%; 145 Control and 145 Affirmed) completed the intervention. That is, 38 students (12%) did not complete the affirmation intervention. 


Condition breakdown among 226 "completers" (students who completed the intervention and both surveys):

```{r}
table(complet226$Intervention)
```
Of the 328 students who consented to participate, 226 (70%; 118 Affirmed and 108 Control) students completed the affirmation exercise as well as both the pre and post-semester questionnaires

## Multiple regression analyses predicting friendship network DVs from intervention condition

### TIME 1

```{r}
# Closeness
modfclose1 <- lm(hclose.T1.Fr ~ Intervention + RecitationC, complet226)
#summary(modfclose1)

# Betweenness
modfbtwn1 <- lm(btwn.T1.Fr ~ Intervention + RecitationC, complet226)
#summary(modfbtwn1)

# Degree (total)
modfdegtot1 <- lm(degtot.T1.Fr ~ Intervention + RecitationC, complet226)
#summary(modfdegtot1)
```

### TIME 2

```{r}
# Closeness
modfclose2 <- lm(hclose.T2.Fr ~ Intervention + RecitationC + hclose.T1.Fr.gmc, complet226) 
#summary(modfclose2)

# Betweenness
modfbtwn2 <- lm(btwn.T2.Fr ~ Intervention + RecitationC + btwn.T1.Fr.gmc, complet226)
#summary(modfbtwn2)

# Degree (total)
modfdegtot2 <- lm(degtot.T2.Fr ~ Intervention + RecitationC + degtot.T1.Fr.gmc, complet226) 
#summary(modfdegtot2)
```

## Permutation tests of network measures

```{r}
# Prep for permutation analyses

numrands <- 200  # set number of permutations to 200 instead of 20,000
fmat1r <- fmat1    # create copy of time 1 network matrix to randomize
fmat2r <- fmat2    # create copy of time 2 network matrix to randomize

complet226sub <- complet226 %>%
  dplyr::select(PPID, Intervention, RecitationC, degtot.T1.Fr.gmc, degtot.T2.Fr,
                degout.T1.Fr.gmc, degout.T2.Fr, degin.T1.Fr.gmc, degin.T2.Fr, hclose.T1.Fr.gmc, 
                hclose.T2.Fr, btwn.T1.Fr.gmc, btwn.T2.Fr, strtotavg.T1.Fr.gmc, strtotavg.T2.Fr,
                stroutavg.T1.Fr.gmc, stroutavg.T2.Fr, strinavg.T1.Fr.gmc, strinavg.T2.Fr)
```

### Time 1 permutation tests
For permutations tests, I copied over the study's original code on how they quantified social networks. Then, similar to what we did in class, stored the p-value of each permutation into a vector, found the percentage of pvalues that that were greater or equal to the observed pvalue, and found the permutation p-value that will be used in Table 2. 

#### Time 1 closeness centrality 
```{r}
fclose1.coef.obs <- modfclose1$coefficients[2] 
# coefficient of effect of intervention condition on closeness centrality
fclose1.coef.out <- vector("numeric")

for(i in 1:numrands){   
  randnodes <- sample(rownames(fmat1), replace = FALSE)   
  rownames(fmat1r) <- randnodes                           
  colnames(fmat1r) <- randnodes                           
   
  # the paper's method of quantifying social networks                                                           
  fg1r <- igraph::graph.adjacency(fmat1r, mode=c("directed"), weighted=TRUE, diag=FALSE)  # form network out of matrix
  close1r <- as.data.frame(harmonic_centrality(fg1r, mode="all", weights=E(fg1r)$weight)/(vcount(fg1r)-1)) # compute closeness centrality of each node in new network
  names(close1r) <- "hclose1r"        # name variable
  close1r$PPID <- row.names(close1r)  # save participant id numbers
  
  complet226subr <- left_join(complet226sub, close1r, by="PPID")  
  
  modfclose1r <- lm(hclose1r ~ Intervention + RecitationC, complet226subr)  # linear model predicting degree in permuted network
  fclose1.coef.out[i] <- modfclose1r$coefficients[2]                      # save intervention coefficient
}

fclose1.p <- sum(abs(fclose1.coef.out) >= abs(fclose1.coef.obs))/numrands
fclose1.p #permutations pvalue 
```

#### Time 1 betweenness centrality

```{r}
fbtwn1.coef.obs <- modfbtwn1$coefficients[2]
fbtwn1.coef.out <- vector("numeric")

for(i in 1:numrands){
  
  randnodes <- sample(rownames(fmat1), replace = FALSE)
  rownames(fmat1r) <- randnodes
  colnames(fmat1r) <- randnodes
  
  fg1r <- igraph::graph.adjacency(fmat1r, mode=c("directed"), weighted=TRUE, diag=FALSE)
  btwn1r <- as.data.frame(betweenness(fg1r, normalized = TRUE))
  names(btwn1r) <- "btwn1r"
  btwn1r$PPID <- row.names(btwn1r)
  complet226subr <- left_join(complet226sub, btwn1r, by="PPID")
  
  modfbtwn1r <- lm(btwn1r ~ Intervention + RecitationC, complet226subr)
  fbtwn1.coef.out[[i]] <- modfbtwn1r$coefficients[2]
  
}

fbtwn1.p <- sum(abs(fbtwn1.coef.out) >= abs(fbtwn1.coef.obs))/numrands
fbtwn1.p 
```

#### Time 1 total degree centrality

```{r}
fdegtot1.coef.obs <- modfdegtot1$coefficients[2]
fdegtot1.coef.out <- vector("numeric")

for(i in 1:numrands){
  
  randnodes <- sample(rownames(fmat1), replace = FALSE)
  rownames(fmat1r) <- randnodes
  colnames(fmat1r) <- randnodes
 
  fg1r <- igraph::graph.adjacency(fmat1r, mode=c("directed"), weighted=TRUE, diag=FALSE)
  
  degtot1r <- as.data.frame(degree(fg1r, mode = "total"))
  names(degtot1r) <- "degtot1r"
  degtot1r$PPID <- row.names(degtot1r)
  
  complet226subr <- left_join(complet226sub, degtot1r, by="PPID")
  
  modfdegtot1r <- lm(degtot1r ~ Intervention + RecitationC, complet226subr)
  fdegtot1.coef.out[[i]] <- modfdegtot1r$coefficients[2]
  
}

fdegtot1.p <- sum(abs(fdegtot1.coef.out) >= abs(fdegtot1.coef.obs))/numrands 
fdegtot1.p
```

### Time 2 permutation tests

#### Time 2 closeness centrality
```{r}
fclose2.coef.obs <- modfclose2$coefficients[2]
fclose2.coef.out <- vector("numeric")

for(i in 1:numrands){
  
  randnodes <- sample(rownames(fmat2), replace = FALSE)
  rownames(fmat2r) <- randnodes
  colnames(fmat2r) <- randnodes
  
  fg2r <- igraph::graph.adjacency(fmat2r, mode=c("directed"), weighted=TRUE, diag=FALSE)
  close2r <- as.data.frame(harmonic_centrality(fg2r, mode="all", weights=E(fg2r)$weight)/(vcount(fg2r)-1))
  names(close2r) <- "hclose2r"
  close2r$PPID <- row.names(close2r)
  complet226subr <- merge(close2r, complet226sub, by="PPID")
  
  modfclose2r <- lm(hclose2r ~ Intervention + RecitationC + hclose.T1.Fr.gmc, complet226subr)
  fclose2.coef.out[[i]] <- modfclose2r$coefficients[2]
  
  #print(i)
}

fclose2.p <- sum(abs(fclose2.coef.out) >= abs(fclose2.coef.obs))/numrands
fclose2.p
```

#### Time 2 betweenness centrality
```{r}
fbtwn2.coef.obs <- modfbtwn2$coefficients[2]
fbtwn2.coef.out <- vector("numeric")

for(i in 1:numrands){
  
  randnodes <- sample(rownames(fmat2), replace = FALSE)
  rownames(fmat2r) <- randnodes
  colnames(fmat2r) <- randnodes
  
  fg2r <- igraph::graph.adjacency(fmat2r, mode=c("directed"), weighted=TRUE, diag=FALSE)
  btwn2r <- as.data.frame(betweenness(fg2r, normalized = TRUE))
  names(btwn2r) <- "btwn2r"
  btwn2r$PPID <- row.names(btwn2r)
  complet226subr <- merge(btwn2r, complet226sub, by="PPID")
  
  modfbtwn2r <- lm(btwn2r ~ Intervention + RecitationC + btwn.T1.Fr.gmc, complet226subr)
  fbtwn2.coef.out[[i]] <- modfbtwn2r$coefficients[2]
  
  # print(i)
}

fbtwn2.p <- sum(abs(fbtwn2.coef.out) >= abs(fbtwn2.coef.obs))/numrands
fbtwn2.p 
```

#### Time 2 total degree centrality and strength of ties
```{r}
fdegtot2.coef.obs <- modfdegtot2$coefficients[2]
fdegtot2.coef.out <- vector("numeric")

for(i in 1:numrands){
  
  randnodes <- sample(rownames(fmat2), replace = FALSE)
  rownames(fmat2r) <- randnodes
  colnames(fmat2r) <- randnodes
  
  fg2r <- igraph::graph.adjacency(fmat2r, mode=c("directed"), weighted=TRUE, diag=FALSE)
  
  degtot2r <- as.data.frame(degree(fg2r, mode = "total"))
  names(degtot2r) <- "degtot2r"
  degtot2r$PPID <- row.names(degtot2r)

  complet226subr <- left_join(complet226sub, degtot2r, by="PPID")
  
  modfdegtot2r <- lm(degtot2r ~ Intervention + RecitationC + degtot.T1.Fr.gmc, complet226subr)
  fdegtot2.coef.out[[i]] <- modfdegtot2r$coefficients[2]

  # print(i)
}

fdegtot2.p <- sum(abs(fdegtot2.coef.out) >= abs(fdegtot2.coef.obs))/numrands
fdegtot2.p
```

## Reproducing Table 2
```{r}
# function to combine all elements to get each row in table 2 
table_2_rows <- function(mod, perm_p) {
  CI <- confint(mod, "InterventionVAffirm", level=.95) # 95% CI for intervention effect
  CI <- paste(round(CI[1],3), "to", round(CI[2],3))
  
  vector <- c(round(mod$coefficients[2],3), #b1obs
  round(summary(mod)$coefficients[2,2],3), #SE
  round(summary(mod)$coefficients[2,3],3), #t
  CI, #Confidence Interval
  round(summary(mod)$coefficients[2,4], 3), #pvalue
  perm_p) #permutation pvalue
  return(vector)
}

#putting each mod into the function to get each row in table 2
t1_closeness <- table_2_rows(modfclose1, fclose1.p)
t1_betweenness <- table_2_rows(modfbtwn1, fbtwn1.p)
t1_degree <- table_2_rows(modfdegtot1, fdegtot1.p)

t2_closeness <- table_2_rows(modfclose2, fclose2.p)
t2_betweenness <- table_2_rows(modfbtwn2, fbtwn2.p)
t2_degree <- table_2_rows(modfdegtot2, fdegtot2.p)

Time1 <- data.frame(rbind(t1_closeness, t1_betweenness, t1_degree))
colnames(Time1) <- c("b_1obs", "SE", "t", "95% CI", "P", "P_perm")

Time2 <- data.frame(rbind(t2_closeness, t2_betweenness, t2_degree))
colnames(Time2) <- c("b_1obs", "SE", "t", "95% CI", "P", "P_perm")

print(Time1)
print(Time2)
```

## Biology track persistence

```{r}
int290 %>% group_by(Intervention) %>% dplyr::summarize(meanpersist = mean(TookSpringClass))
```

Logistic regression:
```{r}
summary(glm(TookSpringClass ~ Intervention + RecitationC, int290, family="binomial"))
```

# EVALUATING THE RESULTS 
## Removing RecitationC Variable 

### TIME 2 regression models without RecitationC variable

```{r}
# Closeness
modclose2 <- lm(hclose.T2.Fr ~ Intervention + hclose.T1.Fr.gmc, complet226) 

# Betweenness

modbtwn2 <- lm(btwn.T2.Fr ~ Intervention + btwn.T1.Fr.gmc, complet226)

# Degree (total)
moddegtot2 <- lm(degtot.T2.Fr ~ Intervention + degtot.T1.Fr.gmc, complet226) 
```


### Reproducing Table 2 without RecitationC
```{r}
t2_closeness2 <- table_2_rows(modclose2, 0)
t2_betweenness2 <- table_2_rows(modbtwn2, 0)
t2_degree2 <- table_2_rows(moddegtot2, 0)

Time2_sansRecitation <- data.frame(rbind(t2_closeness2, t2_betweenness2, t2_degree2)) 
colnames(Time2_sansRecitation) <- c("b_1obs", "SE", "t", "95% CI", "P", "P_perm")
Time2_sansRecitation <- subset(Time2_sansRecitation, select=-c(P_perm))
print(Time2_sansRecitation) #same significance results
```

### Biology track persistence
Logistic regression
```{r}
summary(glm(TookSpringClass ~ Intervention, int290, family="binomial"))
```

## Changing evaluation criteria 
```{r}
#selecting friendship strengths
snat1_interest<- subset(snat1, select= c(PPID, Friend1Strength, Friend2Strength, Friend3Strength, Friend4Strength, Friend5Strength, Friend6Strength))
snat1_interest[is.na(snat1_interest)] <- 0
#summing up friendship strengths
snat1_interest$Friendship <- rowSums(snat1_interest[ , c(2:7)]) 
#selecting participant ID and friendship
t1 <- subset(snat1_interest, select= c(PPID, Friendship))%>%   
  mutate(PPID = as.character(PPID))

snat2_interest<- subset(snat2, select= c(PPID, Friend1Strength, Friend2Strength, Friend3Strength, Friend4Strength, Friend5Strength, Friend6Strength))
snat2_interest[is.na(snat2_interest)] <- 0
snat2_interest$Friendship <- rowSums(snat2_interest[ , c(2:7)])
t2 <- subset(snat2_interest, select= c(PPID, Friendship))%>%
   mutate(PPID = as.character(PPID))

df <- merge(att, t1, by="PPID", all.x=TRUE, all.y = TRUE)
df_all <- merge(df,t2,by="PPID",all.x=TRUE, all.y = TRUE)

df_all <- df_all[complete.cases(df_all), ]
df_all$Difference <- df_all$Friendship.y - df_all$Friendship.x

mod_eval <- lm(Difference ~ Intervention, df_all)
summary(mod_eval)
mod_friendship<- lm(Friendship.y ~ Friendship.x, df_all)
mod_friendship$coefficients
plot(df_all$Friendship.x, df_all$Friendship.y, 
     xlab = "Friendship T1",
     ylab = "Friendship T2")
abline(a = mod_friendship$coefficients[1], b= mod_friendship$coefficients[2], col="blue")
```
This plot shows the linear relationship between Summed Friendship Strenghs in the two time periods. As seen, higher friendship at T1 is associated with higher friendship at T2. 

## Using DAGs - Mediator 
```{r}
#closeness
X <- int290$Intervention 
Z <- int290$hclose.T2.Fr
Y <- int290$TookSpringClass

summary(glm(Y ~ X, family = "binomial")) # x cause y in the overall sense
#but when u add in the mediator z, now it looks like x does not cause y. 
#power of x disappears. 
summary(glm(Y ~ X + Z, family = "binomial")) 

#betweenness
Z <- int290$btwn.T2.Fr

summary(glm(Y ~ X, family = "binomial")) 
summary(glm(Y ~ X + Z, family = "binomial"))

#degree 
Z <- int290$degtot.T2.Fr

summary(glm(Y ~ X, family = "binomial")) 
summary(glm(Y ~ X + Z, family = "binomial"))

dagify(Y ~ Z, Z~X) %>%
  ggdag() + 
  theme_dag()
```


